\section{Dynamic-MSV-Bench and Metrics}
\label{sec:benchmark}

We propose \textbf{Dynamic-MSV-Bench}, comprising 100 high-fidelity multi-shot scenarios across two distinct tracks designed to push current diffusion architectures to their limits.

\subsection{Track A: Semantic Shift (Narrative Diversity)}
Track A evaluates the model's ability to maintain a consistent subject while drastically changing the environment (e.g., from an underwater ruin to a Himalayan peak). 
\textbf{Golden Rule for Track A:} HIGH Background Diversity, HIGH Cut Sharpness, and HIGH Subject Consistency. A low diversity score in this track identifies the \textbf{Static Trap}.

\subsection{Track B: Motion Control (Spatial Integrity)}
Track B focuses on camera dynamics (Pan, Zoom, Tilt) within a fixed environment. This tests the model's spatial reasoning and its ability to follow fine-grained physical instructions.
\textbf{Golden Rule for Track B:} LOW Background Diversity, LOW Cut Sharpness (smooth transitions), and HIGH Subject Consistency. A high diversity score here indicates \textbf{Background Hallucination}.

\subsection{The 4D Decoupled Metrics}
Our framework utilizes four indices to quantify performance:
\begin{enumerate}
    \item \textbf{Subject Consistency ($\mathcal{C}_{subj}$):} DINOv2~\cite{oquab2023dinov2} based similarity of masked subject regions across $K$ shots.
    \item \textbf{Background Diversity ($\mathcal{D}_{bg}$):} Variance of CLIP~\cite{radford2021learning} or DINO embeddings in the environment region.
    \item \textbf{Cut-Transition Sharpness ($\mathcal{S}_{cut}$):} Normalized LPIPS~\cite{zhang2018perceptual} distance peaks at shot boundaries.
    \item \textbf{Diagonal Semantic Alignment (DSA):}
    Our core novelty, DSA, uses column-wise softmax probabilities:
    $$P_{i,j} = \frac{\exp(\tau \cdot M_{i,j})}{\sum_{k=1}^{K} \exp(\tau \cdot M_{k,j})}$$
    $$DSA = \frac{\left( \frac{1}{K} \sum_{i=1}^{K} P_{i,i} \right) - \frac{1}{K}}{1 - \frac{1}{K}}$$
    This formulation ensures that a static video, where all shots are identical, yields a DSA of exactly 0.0.
\end{enumerate}
